PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd
fixed baseline and retro questionnaires

```{r coping edgi gad summary}
if(EDGI == TRUE) {
  summarytools::dfSummary(
    coping.edgi.gad.raw,
    graph.col = F,
    valid.col = F,
    labels.col = F
  )
}
```

@@

  coping.edgi.gad.raw.id <- coping.edgi.gad.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)
    ) %>% 
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping = startDate, # Rename startDate
      gad.problems_made_difficult_care, # Impairment
      gad.pandemic_felt_feelings, # Comparison to before the pandemic (retro)
      gad.feeling_nervous_anxious_or_on_edge_retro = gad.feeling_nervous_anxious_or_on_edge,
      gad.control_worrying_stop_retro = gad.control_worrying_stop,
      gad.worrying_too_much_about_different_things_retro = gad.worrying_too_much_about_different_things,
      gad.trouble_relaxing_retro = gad.trouble_relaxing,
      gad.sit_restless_hard_retro = gad.sit_restless_hard,
      gad.becoming_easily_annoyed_or_irritable_retro = gad.becoming_easily_annoyed_or_irritable,
      gad.awful_feeling_afraid_happen_retro = gad.awful_feeling_afraid_happen,
      gad.feeling_nervous_anxious_or_on_edge_base = gad.feeling_nervous_anxious_or_on_edge.1,
      gad.control_worrying_stop_base = gad.control_worrying_stop.1,
      gad.worrying_too_much_about_different_things_base = gad.worrying_too_much_about_different_things.1,
      gad.trouble_relaxing_base = gad.trouble_relaxing.1,
      gad.sit_restless_hard_base = gad.sit_restless_hard.1,
      gad.becoming_easily_annoyed_or_irritable_base = gad.becoming_easily_annoyed_or_irritable.1,
      gad.awful_feeling_afraid_happen_base = gad.awful_feeling_afraid_happen.1
    ) %>%
    add_numeric(., exclude = exclude_cols_coping_edgi) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r coping edgi pcl cleaning}
if(EDGI == TRUE) {
coping.edgi.pcl.raw.id <- coping.edgi.pcl.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    pcl.stressful_experience_repeated_images_base = pcl.stressful_experience_repeated_images,
    pcl.stressful_experience_upset_reminded_base = pcl.stressful_experience_upset_reminded, # differ to EDGI prepan
    pcl.stressful_situation_avoiding_activities_base = pcl.avoiding_activities_stressful_experience, # differ to EDGI prepan
    pcl.cut_people_feeling_distant_base = pcl.cut_people_feeling_distant,
    pcl.feeling_irritable_or_having_angry_outbursts_base = pcl.feeling_irritable_or_having_angry_outbursts,
    pcl.difficulty_concentrating_base = pcl.difficulty_concentrating,
    pcl.stressful_experience_repeated_images_retro = pcl.stressful_experience_repeated_images.1,
    pcl.stressful_experience_upset_reminded_retro = pcl.stressful_experience_upset_reminded.1,
    pcl.stressful_situation_avoiding_activities_retro = pcl.avoiding_activities_stressful_experience.1,
    pcl.cut_people_feeling_distant_retro = pcl.cut_people_feeling_distant.1,
    pcl.feeling_irritable_or_having_angry_outbursts_retro = pcl.feeling_irritable_or_having_angry_outbursts.1,
    pcl.difficulty_concentrating_retro = pcl.difficulty_concentrating.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

Will need to double check these namings once we have the new cleaned data -- KT. 
```{r coping edgi phq cleaning}
if(EDGI == TRUE) {
coping.edgi.phq.raw.id <- coping.edgi.phq.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)
  ) %>%
  select(
    Sample, # Sample
    ID, # ID
    phq.problems_made_difficult_care,
    phq.pandemic_felt_feelings,
    phq.little_interest_or_pleasure_in_doing_things_retro = phq.little_interest_or_pleasure_in_doing_things,
    phq.feeling_down_depressed_or_hopeless_retro = phq.feeling_down_depressed_or_hopeless,
    phq.staying_asleep_sleeping_trouble_retro = phq.staying_asleep_sleeping_trouble,
    phq.feeling_tired_or_having_little_energy_retro = phq.feeling_tired_or_having_little_energy,
    phq.poor_appetite_or_overeating_retro = phq.poor_appetite_or_overeating,
    phq.feeling_bad_failure_family_retro = phq.feeling_bad_failure_family,
    phq.trouble_concentrating_reading_newspaper_retro = phq.trouble_concentrating_newspaper_reading, # differ EDGI prepandemic
    phq.moving_fidgety_noticed_opposite_retro = phq.moving_fidgety_opposite_slowly, # differ EDGI prepandemic
    phq.dead_hurting_thoughts_retro = phq.dead_hurting_thoughts,
    phq.little_interest_or_pleasure_in_doing_things_base = phq.little_interest_or_pleasure_in_doing_things.1,
    phq.feeling_down_depressed_or_hopeless_base = phq.feeling_down_depressed_or_hopeless.1,
    phq.staying_asleep_sleeping_trouble_base = phq.staying_asleep_sleeping_trouble.1,
    phq.feeling_tired_or_having_little_energy_base = phq.feeling_tired_or_having_little_energy.1,
    phq.poor_appetite_or_overeating_base = phq.poor_appetite_or_overeating.1,
    phq.feeling_bad_failure_family_base = phq.feeling_bad_failure_family.1,
    phq.trouble_concentrating_reading_newspaper_base = phq.trouble_concentrating_newspaper_reading.1, # differ EDGI prepandemic
    phq.moving_fidgety_noticed_opposite_base = phq.moving_fidgety_opposite_slowly.1, # differ EDGI prepandemic
    phq.dead_hurting_thoughts_base = phq.dead_hurting_thoughts.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

    ocir.repeat_feel_numbers_retro = ocir.numbers_repeat_feel, ## Different from EDGI prepandemic

@@

@@

    ocir.control_difficult_find_thoughts_retro = ocir.control_find_difficult_thoughts, ## Different from EDGI prepandemic

@@

    ocir.repeat_feel_numbers_base = ocir.numbers_repeat_feel.1, ## Different from EDGI prepandemic

@@

@@

    ocir.control_difficult_find_thoughts_base = ocir.control_find_difficult_thoughts.1, ## Different from EDGI prepandemic

>>>>

```{r coping edgi gad summary}
if(EDGI == TRUE) {
  summarytools::dfSummary(
    coping.edgi.gad.raw,
    graph.col = F,
    valid.col = F,
    labels.col = T # to see what is baseline and what is retro
  )
}
```

@@

  coping.edgi.gad.raw.id <- coping.edgi.gad.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)
    ) %>% 
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping = startDate, # Rename startDate
      gad.problems_made_difficult_care = gad.care_difficult_home_things, # Impairment (different to GLAD)
      gad.pandemic_felt_feelings = gad.pandemic_feelings_felt, # Comparison to before the pandemic (retro; different to GLAD)
      gad.feeling_nervous_anxious_or_on_edge_retro = gad.feeling_nervous_anxious_or_on_edge,
      gad.control_worrying_stop_retro = gad.control_worrying_stop,
      gad.worrying_too_much_about_different_things_retro = gad.worrying_too_much_about_different_things,
      gad.trouble_relaxing_retro = gad.trouble_relaxing,
      gad.sit_restless_hard_retro = gad.sit_restless_hard,
      gad.becoming_easily_annoyed_or_irritable_retro = gad.becoming_easily_annoyed_or_irritable,
      gad.awful_feeling_afraid_happen_retro = gad.feeling_afraid_awful_happen, # different to GLAD: gad.awful_feeling_afraid_happen,
      gad.feeling_nervous_anxious_or_on_edge_base = gad.feeling_nervous_anxious_or_on_edge.1,
      gad.control_worrying_stop_base = gad.control_worrying_stop.1,
      gad.worrying_too_much_about_different_things_base = gad.worrying_too_much_about_different_things.1,
      gad.trouble_relaxing_base = gad.trouble_relaxing.1,
      gad.sit_restless_hard_base = gad.sit_restless_hard.1,
      gad.becoming_easily_annoyed_or_irritable_base = gad.becoming_easily_annoyed_or_irritable.1,
      gad.awful_feeling_afraid_happen_base = gad.feeling_afraid_awful_happen.1 # different to GLAD: gad.awful_feeling_afraid_happen.1
    ) %>%
    add_numeric(., exclude = exclude_cols_coping_edgi) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r coping edgi pcl cleaning}
if(EDGI == TRUE) {
coping.edgi.pcl.raw.id <- coping.edgi.pcl.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    pcl.stressful_experience_repeated_images_base = pcl.repeated_images_thoughts_past, # different pcl.stressful_experience_repeated_images,
    pcl.stressful_experience_upset_reminded_base = pcl.upset_reminded_past_feeling, # differ to EDGI prepan
    pcl.stressful_situation_avoiding_activities_base = pcl.avoiding_activities_reminded_situations, # differ to EDGI prepan
    pcl.cut_people_feeling_distant_base = pcl.cut_people_feeling_distant,
    pcl.feeling_irritable_or_having_angry_outbursts_base = pcl.feeling_irritable_or_having_angry_outbursts,
    pcl.difficulty_concentrating_base = pcl.difficulty_concentrating,
    pcl.stressful_experience_repeated_images_retro = pcl.repeated_images_thoughts_past.1,
    pcl.stressful_experience_upset_reminded_retro = pcl.upset_reminded_past_feeling.1,
    pcl.stressful_situation_avoiding_activities_retro = pcl.avoiding_activities_reminded_situations.1,
    pcl.cut_people_feeling_distant_retro = pcl.cut_people_feeling_distant.1,
    pcl.feeling_irritable_or_having_angry_outbursts_retro = pcl.feeling_irritable_or_having_angry_outbursts.1,
    pcl.difficulty_concentrating_retro = pcl.difficulty_concentrating.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r coping edgi phq cleaning}
if(EDGI == TRUE) {
coping.edgi.phq.raw.id <- coping.edgi.phq.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)
  ) %>%
  select(
    Sample, # Sample
    ID, # ID
    phq.problems_made_difficult_care = phq.care_difficult_home_things,
    phq.pandemic_felt_feelings = phq.pandemic_feelings_felt,
    phq.little_interest_or_pleasure_in_doing_things_retro = phq.little_interest_or_pleasure_in_doing_things,
    phq.feeling_down_depressed_or_hopeless_retro = phq.feeling_down_depressed_or_hopeless,
    phq.staying_asleep_sleeping_trouble_retro = phq.staying_asleep_sleeping_trouble,
    phq.feeling_tired_or_having_little_energy_retro = phq.feeling_tired_or_having_little_energy,
    phq.poor_appetite_or_overeating_retro = phq.poor_appetite_or_overeating,
    phq.feeling_bad_failure_family_retro = phq.family_feeling_bad_failure, # differ GLAD phq.feeling_bad_failure_family,
    phq.trouble_concentrating_reading_newspaper_retro = phq.trouble_concentrating_newspaper_reading, # differ EDGI prepandemic
    phq.moving_fidgety_noticed_opposite_retro = phq.moving_fidgety_opposite_slowly, # differ EDGI prepandemic
    phq.dead_hurting_thoughts_retro = phq.dead_hurting_thoughts,
    phq.little_interest_or_pleasure_in_doing_things_base = phq.little_interest_or_pleasure_in_doing_things.1,
    phq.feeling_down_depressed_or_hopeless_base = phq.feeling_down_depressed_or_hopeless.1,
    phq.staying_asleep_sleeping_trouble_base = phq.staying_asleep_sleeping_trouble.1,
    phq.feeling_tired_or_having_little_energy_base = phq.feeling_tired_or_having_little_energy.1,
    phq.poor_appetite_or_overeating_base = phq.poor_appetite_or_overeating.1,
    phq.feeling_bad_failure_family_base = phq.family_feeling_bad_failure.1, # differ GLAD
    phq.trouble_concentrating_reading_newspaper_base = phq.trouble_concentrating_newspaper_reading.1, # differ EDGI prepandemic
    phq.moving_fidgety_noticed_opposite_base = phq.moving_fidgety_opposite_slowly.1, # differ EDGI prepandemic
    phq.dead_hurting_thoughts_base = phq.dead_hurting_thoughts.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

    ocir.repeat_feel_numbers_retro = ocir.repeat_numbers_feel, # Different from EDGI prepandemic ocir.numbers_repeat_feel,

@@

    ocir.light_switches_water_taps_retro = ocir.water_taps_light_switches, #  differ ocir.light_switches_water_taps,

@@

    ocir.control_difficult_find_thoughts_retro = ocir.control_difficult_thoughts_find, # Differ to EDGI prepan ocir.control_find_difficult_thoughts,

@@

    ocir.repeat_feel_numbers_base = ocir.repeat_numbers_feel.1, ## Different from EDGI prepandemic

@@

    ocir.light_switches_water_taps_base = ocir.water_taps_light_switches.1, # differ

@@

    ocir.control_difficult_find_thoughts_base = ocir.control_difficult_thoughts_find.1, ## Different from EDGI prepandemic
